---
title: "Covid91 vaccine study FinalFall2022"
author: "Valerie Rodriguez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2023" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Coding details







## Males

###Barplot

```{r}
barchartGC(~infected + treatment,data=guy)
barchartGC(~infected + treatment,data=guy, type="percent")

```

This bar plot represents males relationship between treatment and whether or not they were infected with Covid-19. In the percent graph we can see that close to 70% of placebo participants got Covid compared to about 30% who were vaccinated and got Covid. Those who were vaccinated and those who got a placebo are both at about 50% in regards to being "OK" (not getting Covid). In the frequency graph both groups had relatively the same frequency of being "OK" (not contracting Covid). While those who got a placebo were slightly higher in regards to getting Covid which was at about 300. In contrast those who were vaccinated were about about 100 in regards to getting Covid.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

Through the numerical results we can see that 68.44% of placebo participants got Covid while only 31.56% of vaccine participants got Covid. Those who were vaccinated and got Covid are less than half of those who got the placebo and still got infected. While 49.88% of placebo participants didn't get Covid and 50.12% of vaccine participants didn't get Covid. These numbers are very close like seen and expected in the bar plot.

In regards to those who received the placebo only 2.69% got Covid while 97.31% were "OK" (not infected). Those who got the vaccine only had 1.25% contraction rate of Covid while 98.75% were "OK". The difference between whether or not you got vaccinated in these statistics doesn't seem to have really any much effect on whether or not you contract Covid.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

In these inferential results the P value is 2.364e-12 which is a very small number. This means that these results are not likely to occur at random.

The observed results show that there is a significant difference between whether or not males are getting Covid if they are vaccinated or not. 245 of those who got the placebo got Covid while only 113 males who got vaccinated got Covid. The expected results suggested that these numbers would be close but in reality they aren't. This aligns with the previous data we saw being that the vaccine does have a positive affect on males and lowers their chances of getting Covid. The odds ratio 2.17 is right in between the confidence interval 1.73-2.75










## Females




## LGBTQ




## Druggies




# Overall Results and Conclusions


