---
title: "Covid91 vaccine study FinalFall2022"
author: "Valerie Rodriguez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2023" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Coding details

I changed the character variable from FinalData to factor Variables by using :
FinalData<- FinalData %>%
mutate(across(where(is.character), as.factor))
Then I had to make sure that all the subcategories of guys, gals, gays and drug user separate so that they could be assessed individually. I stored them as separate values and this can be seen when the assignment is knitted. By separating them I was able to see if the vaccine was effective for each subcategory.


## Males

###Barplot

```{r}
barchartGC(~infected + treatment,data=guy)
barchartGC(~infected + treatment,data=guy, type="percent")

```

This bar plot represents males relationship between treatment and whether or not they were infected with Covid-19. In the percent graph we can see that close to 70% of placebo participants got Covid compared to about 30% who were vaccinated and got Covid. Those who were vaccinated and those who got a placebo are both at about 50% in regards to being "OK" (not getting Covid). In the frequency graph both groups had relatively the same frequency of being "OK" (not contracting Covid). While those who got a placebo were slightly higher in regards to getting Covid which was at about 300. In contrast those who were vaccinated were about about 100 in regards to getting Covid.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

Through the numerical results we can see that 68.44% of placebo participants got Covid while only 31.56% of vaccine participants got Covid. Those who were vaccinated and got Covid are less than half of those who got the placebo and still got infected. While 49.88% of placebo participants didn't get Covid and 50.12% of vaccine participants didn't get Covid. These numbers are very close like seen and expected in the bar plot.

In regards to those who received the placebo only 2.69% got Covid while 97.31% were "OK" (not infected). Those who got the vaccine only had 1.25% contraction rate of Covid while 98.75% were "OK". The difference between whether or not you got vaccinated in these statistics doesn't seem to have really any much effect on whether or not you contract Covid.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

In these inferential results the P value is 2.364e-12 which is a very small number. This means that these results are not likely to occur at random.

The observed results show that there is a significant difference between whether or not males are getting Covid if they are vaccinated or not. 245 of those who got the placebo got Covid while only 113 males who got vaccinated got Covid. The expected results suggested that these numbers would be close but in reality they aren't. This aligns with the previous data we saw being that the vaccine does have a positive affect on males and lowers their chances of getting Covid. The odds ratio 2.17 is right in between the confidence interval 1.73-2.75.


```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The data above shows the effectiveness for the vaccine for guys. We know that if the effectiveness is above 50% then the vaccine can be deemed effective. Since out interval is 63.4 to 73.35 we can say that the vaccine is effective for guys.


## Females


###Barplot

```{r}
barchartGC(~infected + treatment,data=gal)
barchartGC(~infected + treatment,data=gal, type="percent")

```

From the percent data we can see that gals had a higher chance of getting Covid if vaccinated in compared to the women who received the placebo. About 56% of vaccinated women got Covid while about 45% of placebo participates got Covid. The percent and frequency of women who were okay was relatively the same regardless if they were vaccinated or not. The frequency of women who were vaccinated and got Covid was slightly larger (about 200) than those who got the placebo (about 100).


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

From the numerical results we can see that women who were vaccinated and got Covid was 57 compared to 50 who were "OK". Women who got the placebo and still got Covid was 43 compared to 50 who were "OK". We can see that those women who got vaccinated, it actually ended up hurting them because they contracted Covid more. We can see this being proven again in the lower half of data results. Women who had the placebo got Covid 1.7% while women who were vaccinated got Covid 2.3%.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```

From the inferential results the p value is 0.005 which is small meaning that the expected results are very unlikely to happen.

The observed results show us that 155 women who got the placebo ended up getting Covid while 207 women who were vaccinated got Covid. The expected results were supposed to be an even amount of both sides were going to contract Covid. The odds ratio in the Fisher test is 0.74 which is almost exactly in between the confidence interval which is 0.60-0.92. This all aligns with the data from the pervious charts.

```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

In regards to the effectiveness of a vaccine we want our confidence interval to be over 50% for the vaccine to be deemed effective. Yet the women's confidence interval is from 37.34 to 47.86 which means that this vaccine is not effective for women.



## LGBTQ

### Barplot

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")

```

From the percent bar plot we can see that gays that got vaccinated have gotten Covid significantly more than those who got the placebo. About 90% of vaccinated gays got Covid while about 15% on placebo gays got Covid. In regards to those who were "OK" and vaccinated were slightly more than those who got the placebo. The frequency at which placebo gays got Covid is very small compared to those who were vaccinated. In regards to those who were "OK" the vaccinated people were slightly higher than those who were not.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

For this data we can see that 87 vaccinated gays got Covid and 51 did not. While only 13 placebo gays got Covid and 49 did not. Those who were vaccinated were much more likely to get Covid. 3.3% out of 96.7% of vaccinated participants got Covid. 0.54% out of 99.46% of placebo participants didn't get Covid.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```

In these inferential results the p value is 1e-08 which is very small number. This means that the expected results of 29 out of 1442 placebo and 31 out of 1558 vaccinated people getting Covid is unlikely to happen. The observed counts shows that 8 out 1463 placebo participates got Covid while 52 out of 1537 vaccinated people got Covid. The odds ratio in the Fisher test is 0.16 which isn't as in the middle of the confidence interval as the previous subcategories.


```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

If the effectiveness confidence interval is above 50% then we know that the vaccine is effective for this group. For those participants that are gay the effectiveness of the vaccine is 6.2 to 25.61 which is extremely low. This vaccine is ineffective for gays and this aligns with the previous data we saw.


## Druggies

### Barplot

```{r}
barchartGC(~infected + treatment,data=Druguser)
barchartGC(~infected + treatment,data=Druguser, type="percent")
```

In the percent plot the drug users that are vaccinated got Covid significantly more than those who got the placebo. About 85% of vaccinated drug uses got Covid while about 18% of placebo drug users got Covid. Those who were "OK" are relatively close in regards to whether or not they got the vaccine or not (vaccinated slightly more). In the frequency plot those who were vaccinated are 3-4x more likely to get Covid compared to placebo participants. Vaccinated people are slightly more "OK" compared to placebo drug users.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```

According to the numerical results we are seeing the same trend. In regards to the placebo drug users 17 got Covid while 49 are "OK". In regards to the vaccinated drug users 83 got Covid while 51 were "OK".  Placebo vaccine drug users got Covid 0.81% of the time compared to  99.19% were "OK". Vaccinated drug users got Covid 3.8% of the time compared to 96.2% that were "OK". We are continuing to see that vaccinated drug users are more likely to get Covid compared to placebo participants.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```


In these inferential results the p value is 3e-07 which is very small number. This means that the expected results of 29 out of 1206 placebo and 31 out of 1294 vaccinated people getting Covid is unlikely to happen. The observed counts shows that 10 out 1225 placebo participates got Covid while 50 out of 1275 vaccinated people got Covid. The odds ratio in the Fisher test is 0.21 which is in the middle of the confidence interval as the first few subcategories.


```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

For this vaccine to be effective for drug users the effectiveness confidence interval must be over 50%. For drug users the effectiveness of the vaccine is 8.57 to 29.47 which means that it is not effective. 

# Overall Results and Conclusions

After analyzing all of the data for guys, gals, gays and drug users only guys have seen effectiveness in getting the vaccine and not getting Covid. The vaccine is not effective for gals, gays and drug users. Below I have ranked each group and the vaccine effectiveness.

Vaccine effectiveness:
1) Guys 63.4 to 73.35
2) Gals 37.34 to 47.86
3) Drug user 8.57 to 29.47
4) Gays 6.2 to 25.61
