---
title: "Covid91 vaccine study FinalFall2022"
author: "Valerie Rodriguez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2023" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

###Barplot

```{r}
barchartGC(~infected + treatment,data=guy)
barchartGC(~infected + treatment,data=guy, type="percent")

```

This bar plot represents males relationship between treatment and whether or not they were infected with Covid-19. In the percent graph we can see that close to 70% of placebo participants got Covid compared to about 30% who were vaccinated and got Covid. Those who were vaccinated and those who got a placebo are both at about 50% in regards to being "OK" (not getting Covid). In the frequency graph both groups had relatively the same frequency of being "OK" (not contracting Covid). While those who got a placebo were slightly higher in regards to getting Covid which was at about 300. In contrast those who were vaccinated were about about 100 in regards to getting Covid.

## Females




## LGBTQ




## Druggies




# Overall Results and Conclusions


